---
title: "PeerAssessment1"
author: "me"
date: "12 de enero de 2015"
output: html_document
---


## Loading and preprocessing the data

```{r, echo=TRUE}

library(plyr)
activity <- read.csv("activity.csv", stringsAsFactors=FALSE)
act_day<-ddply(activity, .(date), summarize, total_steps=sum(steps))

```

## What is mean total number of steps taken per day?

```{r, echo=TRUE, results='asis'}

hist(act_day[,2], main="Histogram of the total number of steps taken each day", col="green", xlab="Number of steps")
mean_step<-mean(act_day[,2], na.rm=TRUE)
median_step<-median(act_day[,2], na.rm=TRUE)

rep<-rbind(mean_step, median_step)
colnames(rep)<-c("Steps per day")
rownames(rep)<-c("Mean", "Median")

library(xtable)
xt<-xtable(rep)
print(xt, type="html")
```

## What is the average daily activity pattern?

```{r, echo=TRUE}

act_inter<-ddply(activity, .(interval), summarize, total_steps=mean(steps, na.rm=TRUE))
plot(act_inter, type="l", col="blue", ylab="Number of steps")

```

```{r, echo=TRUE}

maxim<-act_inter[which(act_inter[,2]==max(act_inter[,2])),1]
```
The 5-minute interval which contains the maximum number of steps, on average across all the days in the dataset is `r maxim`.


## Imputing missing values

```{r, echo=TRUE}

na_tot<-dim(activity)[1]-dim(na.omit(activity))[1]
```


The total number of rows with missing values is `r na_tot`.

```{r, echo=TRUE, results='asis'}
#Imputing missing values by using median for each interval

act_inter_med<-ddply(activity, .(interval), summarize, total_steps=median(steps, na.rm=TRUE))
activity_full<-activity
for(i in 1:dim(act_inter_med)[1])
        {
                activity_full[which((is.na(activity_full[,1]))&(activity_full[,3]==act_inter_med[i,1])),1]<-act_inter_med[i,2]
        }

act_day_full<-ddply(activity_full, .(date), summarize, total_steps=sum(steps))

hist(act_day_full[,2], main="Histogram of the total number of steps taken each day (imputed data)", col="red", xlab="Number of steps")
mean_stepf<-mean(act_day_full[,2], na.rm=TRUE)
median_stepf<-median(act_day_full[,2], na.rm=TRUE)

repf<-rbind(mean_stepf, median_stepf)
colnames(repf)<-c("Steps per day")
rownames(repf)<-c("Mean", "Median")

library(xtable)
xtf<-xtable(repf)
print(xtf, type="html")

```
The mean and median differs from the estimates from the first part of the assignment.
Imputing missing data makes the differences between the mean and median larger.

```{r, echo=TRUE, results='asis'}
repfim<-cbind(rep, repf)
colnames(repfim)<-c("No imputed", "Imputed")
xtfim<-xtable(repfim)
print(xtfim, type="html")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
##Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

activity_full<-mutate(activity_full, weekday=weekdays(as.Date(activity_full[,2])))
activity_full[which(activity_full[,4] %in% c("lunes","martes","miércoles", "jueves", "viernes")),4]<-"weekday"
activity_full[which(activity_full[,4] %in% c("sábado","domingo")),4]<-"weekend"

library(ggplot2)

act_week_full<-ddply(activity_full, .(interval, weekday), summarize, total_steps=sum(steps))

qplot(interval, total_steps, data=act_week_full, geom=c("line"), facets=.~weekday, main="Differences in activity patterns between weekdays and weekends")        

```

